---
title: "Quasi-Poisson Policy Analysis of Workplace Respiratory Infections"
author: "Truong Lieu Dong Dang"
date: "2026-07-18"
output:
  html_document:
    toc: true
    toc_depth: 3
    theme: readable
  pdf_document:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center"
)
```

## Executive Summary

This project analyses aggregated workplace respiratory infection data from a hospital network. The objective is to identify which operational factors are associated with lower infection rates after accounting for different levels of work exposure.

The response variable is a count of infections, so the analysis starts with a Poisson generalised linear model. Because each record represents a different number of worked hours, the model includes `log(hours)` as an offset. A goodness-of-fit check shows severe overdispersion, so the final inference uses a Quasi-Poisson model with corrected standard errors.

The final model indicates that nursing experience is the dominant predictor. Compared with Experience Level 1, Experience Levels 2, 3, and 4 are associated with substantially lower infection rates. PPE regime, wage-increase coverage, and external training coverage do not provide meaningful additional explanatory value after experience is included.

## Data Preparation

```{r data-preparation}
library(tidyverse)
library(ggpubr)

infections <- read.csv("../data/infections.csv", header = TRUE)

infections <- infections %>%
  mutate(
    ppe_regime = factor(ppe_regime, levels = 1:4),
    experience = factor(experience, levels = 1:4),
    infection_rate = infections / hours * 1000
  )

summary(infections)
sum(is.na(infections))
```

The derived `infection_rate` variable represents infections per 1,000 worked hours. It is used for descriptive visualisation, while the count model uses raw infection counts with an exposure offset.

## Exploratory Analysis

```{r eda-plots, fig.width=10, fig.height=8}
plot_ppe <- ggplot(infections, aes(x = ppe_regime, y = infection_rate, fill = ppe_regime)) +
  geom_boxplot(alpha = 0.75, outlier.colour = "#B42318", outlier.shape = 16) +
  labs(
    title = "Infection Rate by PPE Regime",
    x = "PPE regime",
    y = "Infections per 1,000 hours"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

plot_experience <- ggplot(infections, aes(x = experience, y = infection_rate, fill = experience)) +
  geom_boxplot(alpha = 0.75, outlier.colour = "#B42318", outlier.shape = 16) +
  labs(
    title = "Infection Rate by Experience Level",
    x = "Experience level",
    y = "Infections per 1,000 hours"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

plot_wage <- ggplot(infections, aes(x = wage_inc, y = infection_rate)) +
  geom_point(colour = "#1F4E79", alpha = 0.65) +
  geom_smooth(method = "lm", se = TRUE, colour = "#B42318", linetype = "dashed") +
  labs(
    title = "Wage-Increase Coverage",
    x = "Proportion receiving wage increase",
    y = "Infections per 1,000 hours"
  ) +
  theme_minimal()

plot_training <- ggplot(infections, aes(x = training, y = infection_rate)) +
  geom_point(colour = "#246B45", alpha = 0.65) +
  geom_smooth(method = "lm", se = TRUE, colour = "#B42318", linetype = "dashed") +
  labs(
    title = "External Training Coverage",
    x = "Proportion completing external training",
    y = "Infections per 1,000 hours"
  ) +
  theme_minimal()

ggarrange(plot_ppe, plot_experience, plot_wage, plot_training, ncol = 2, nrow = 2)
```

The descriptive plots suggest that infection rates decline as experience increases. PPE regime appears to differ in the raw plots, but this pattern must be tested in a multivariable model because PPE allocation may be correlated with experience mix. Wage-increase and training coverage do not show strong linear relationships with infection rates in the exploratory plots.

## Modelling Strategy

The initial Poisson model is:

```{r poisson-model}
poisson_model <- glm(
  infections ~ ppe_regime + experience + wage_inc + training,
  offset = log(hours),
  family = poisson(link = "log"),
  data = infections
)

summary(poisson_model)

gof_p_value <- pchisq(
  poisson_model$deviance,
  df = poisson_model$df.residual,
  lower.tail = FALSE
)

dispersion_ratio <- poisson_model$deviance / poisson_model$df.residual
overdispersion_threshold <- 1 + 3 * sqrt(2 / poisson_model$df.residual)

tibble(
  goodness_of_fit_p_value = gof_p_value,
  dispersion_ratio = dispersion_ratio,
  overdispersion_threshold = overdispersion_threshold,
  overdispersed = dispersion_ratio > overdispersion_threshold
)
```

The dispersion ratio is well above the overdispersion threshold, so standard Poisson uncertainty is too narrow. The final inference therefore uses a Quasi-Poisson model.

```{r quasipoisson-full}
full_quasi_model <- glm(
  infections ~ ppe_regime + experience + wage_inc + training,
  offset = log(hours),
  family = quasipoisson(link = "log"),
  data = infections
)

summary(full_quasi_model)
drop1(full_quasi_model, test = "F")
```

## Covariate Selection

Backward elimination is performed with F-tests because quasi-likelihood models do not produce a standard AIC.

```{r model-selection}
without_training <- update(full_quasi_model, . ~ . - training)
anova(without_training, full_quasi_model, test = "F")

without_ppe <- update(without_training, . ~ . - ppe_regime)
anova(without_ppe, without_training, test = "F")

final_model <- update(without_ppe, . ~ . - wage_inc)
anova(final_model, without_ppe, test = "F")

summary(final_model)
```

After accounting for experience, the other covariates do not improve model fit enough to justify retaining them. The final model is:

```{r final-model-irr}
coef_table <- cbind(
  estimate = coef(final_model),
  confint.default(final_model)
)

irr_table <- round(exp(coef_table), 3)
irr_table
```

## Interpretation

The final model estimates infection-rate ratios relative to Experience Level 1:

- Experience Level 2: approximately 58% lower infection rate.
- Experience Level 3: approximately 57% lower infection rate.
- Experience Level 4: approximately 75% lower infection rate.

These results support a practical policy interpretation: workforce experience and retention should be treated as central infection-control assets, not merely staffing variables.

## Diagnostics

```{r diagnostics, fig.width=10, fig.height=8}
par(mfrow = c(2, 2))
plot(final_model)
par(mfrow = c(1, 1))

standardised_residuals <- rstandard(final_model)
which(abs(standardised_residuals) > 1.96)
```

The final diagnostic plots should be read in the context of count data and substantial overdispersion. The Quasi-Poisson model corrects the main uncertainty issue, and no single operational recommendation should be made from an individual observation alone.

## Recommendations

1. Prioritise retention of experienced nursing staff as an infection-control strategy.
2. Pair less experienced nurses with senior mentors in high-risk units.
3. Review PPE policy as part of a broader system, rather than selecting one PPE regime based only on raw infection rates.
4. Reassess external training content and uptake measurement, since training coverage alone does not explain lower infection rates in this dataset.

## Limitations

This is an observational analysis of aggregated records. The findings should not be interpreted as individual-level causality. Important unmeasured factors, such as patient load, ward acuity, shift patterns, and local outbreak intensity, may also influence infection rates.
